---
output:
  html_document:
    theme: yeti
    highlight: espresso
    toc: true
    toc_float: false
    toc_depth: 2
---




<!-- Javascript for zooming on figures (adapted from: https://stackoverflow.com/questions/40401680) -->

<!-- Jquery import conflicts with DT::datatable so needs to be commented here -->
<!-- <script src="https://ajax.googleapis.com/ajax/libs/jquery/3.1.1/jquery.min.js"></script> -->

<style>
.zoomDiv {
  display: none;
  position: fixed;
  top: 50%;
  left: 50%;
  z-index: 50;
  transform: translate(-50%, -50%);
  background-color: #FFFFFF;
  box-shadow: 0px 0px 50px #888888;
  width: fit-content;
  max-width: 90%;
  max-height: 90%;
  overflow: auto;
}

.zoomImg {
  width: 100%;
}
</style>

<script type="text/javascript">
  $(document).ready(function() {
    $('body').prepend("<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>");
    // onClick for all img except the zoomed one and link ones (filter)
    // use 'img.zoom' and out.extra='class=\"zoom\"' in chunk to specify manually which chunk images can be zoomed
    $('img:not(.zoomImg)').filter(':not(a *)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src'));
      $('.zoomDiv').show();
    });
    // onClick function for hiding div
    $('img.zoomImg').click(function() {
      $('.zoomDiv').hide();
    });
  });
</script>




```{r rmd_setKnitrOptions, include=FALSE}
# Set hook for PNG optimization ('optipng' must be installed on system)
#knitr::knit_hooks$set(optipng = knitr::hook_optipng)

# Set chunks default options
knitr::opts_chunk$set(results='asis',
                      echo = FALSE,           # Do not show chunks code
                      message = FALSE,        # Hide R regular text output in reports
                      warning = FALSE,        # Show warning messages
                      error = TRUE,           # Do not stop generating report on errors
                      cache = FALSE,          # Disable chunk result caching
                      cache.comments = FALSE, # Changing comments does not invalidate cache
                      fig.dim = c(5, 5),      # Size of device for R figures
                      autodep = TRUE,         # Automatic chunk dependencies if cache enabled
                      optipng = '-o2');       # PNG optimization level (2=default, 5=slow, 7=very slow)
```




```{r rmd_loadParamsAndDependencies, warning=FALSE, cache = FALSE}
### Define working folder
WORKING_DIR   = getwd();


### Load libraries, global functions/variables, and chunks
source( file.path( WORKING_DIR, "00_generalDeps.R"));

# Import chunks from source files
read_chunk( path=file.path( WORKING_DIR, "01_prepareData.R"));
read_chunk( path=file.path( WORKING_DIR, "02_variableGenes.R"));
read_chunk( path=file.path( WORKING_DIR, "03_cellHeterogeneity_PCA.R"));
read_chunk( path=file.path( WORKING_DIR, "04_cellHeterogeneity_tSNE_UMAP.R"));
read_chunk( path=file.path( WORKING_DIR, "05_cellHeterogeneity_Clustering.R"));
read_chunk( path=file.path( WORKING_DIR, "06_cellHeterogeneity_MarkerGenes.R"));
read_chunk( path=file.path( WORKING_DIR, "90_finalize.R"));
```




---
title: `r paste( GLOBAL_DESCRIPTION, ANALYSIS_STEP_NAME, sep ="<BR>")`
---




```{r rmd_initFunctions}

### Create output directory
dir.create( PATH_ANALYSIS_OUTPUT, showWarnings = FALSE);
PATH_ANALYSIS_EXTRA_OUTPUT = file.path( PATH_ANALYSIS_OUTPUT, "Extra")
dir.create( PATH_ANALYSIS_EXTRA_OUTPUT, showWarnings = FALSE);


### Record starting time (for report and file names)
startTime = Sys.time();
startTimeReport   = format( startTime, "%Y %m %d %H:%M:%S")
startTimeFileName = format( startTime, "%Y_%m_%d_%H_%M_%S")


### Initialize utils

# Create a hash for all parameters
paramsHash = digest( environment(), algo="murmur32");

# Set seed for pseudo random number generator
set.seed(SEED);

# Set the use of multiple cores using 'future' library, as implemented in Seurat3 when possible
# https://github.com/satijalab/seurat/issues/457
plan(strategy = "multicore", workers = NBCORES);

```




# Primary analysis (QC)




## Run informations {.tabset .tabset-fade}

### Run
<center>
```{r rmd_runInfo_run}
pander( c( 'Date'     = startTimeReport,
           'Sample'   = EXPERIMENT_NAME,
           'Hash'     = paramsHash));
```
</center>

### Parameters
```{r rmd_runInfo_params}
listParams = as.list( paramsEnv);
pander(listParams[order(names(listParams))]); # Show params in alphabetical order
```

### System
<center>
```{r rmd_runInfo_sysInfo}
showSimpleDT( Sys.info())
```
</center>

### R
```{r rmd_runInfo_r}
pander( sessionInfo());
```

<!-- Ending the tabset section -->
<!-- ## {.toc-ignore} -->
<br /><br />


## Load Data

```{r rmd_dataLoad, warning = TRUE, fig.dim = c(9,6)}
<<loadData>>
```


## Filtering cells

```{r rmd_dataFilterAndNormalize, out.width='50%'}
<<filterData_selection>>
<<filterData_summaryPlot>>
<<filterData_filterObject>>
<<normalizeData>>
```
<br /><br />




## Variable genes
```{r rmd_findVariableGenes}
<<findVariableGenes_seuratMethod>>
```

Number of variable genes found: `r variablesGenesStats`

```{r rmd_findVariableGenes_summaryPlot}
<<findVariableGenes_summaryPlot>>
```
<br />

Top `r VARIABLE_FEATURES_SHOWTOP` most variable annotations:

```{r rmd_findVariableGenes_summaryTable}
<<findVariableGenes_summaryTable>>
```
<br /><br />





## PCA results

PCA projections for dimensions `r paste( 1:PCA_PLOTS_NBDIMS, collapse="/")` (colored by batch)  
```{r rmd_heterogeneity_pca, out.width='33%'}
<<heterogeneity_pca>>
```
<br />

UMIs counts on PCA dimensions `r paste( 1:PCA_PLOTS_NBDIMS, collapse="/")`  
```{r rmd_heterogeneity_pca_umisCounts, out.width='33%'}
<<heterogeneity_pca_umisCounts>>
```
<br />

Genes counts on PCA dimensions `r paste( 1:PCA_PLOTS_NBDIMS, collapse="/")`  
```{r rmd_heterogeneity_pca_genesCounts, out.width='33%'}
<<heterogeneity_pca_genesCounts>>
```
<br />

Correlation (Spearman) of PCA dimensions `r paste( 1:PCA_PLOTS_NBDIMS, collapse="/")` with UMIs and genes counts  
```{r rmd_heterogeneity_pca_correlations, fig.dim = c(9,6)}
<<heterogeneity_pca_correlations>>
```
<br />

PCA loadings of top `r PCA_PLOTS_NBFEATURES` features for dimensions `r paste( 1:PCA_PLOTS_NBDIMS, collapse="/")`  
```{r rmd_heterogeneity_pca_loadings, fig.dim = c(4, 4), out.width='33%'}
<<heterogeneity_pca_loadings>>
```
<br /><br />




## Dimensionality reduction
```{r rmd_heterogeneity_dimReduc, warning=FALSE}
<<heterogeneity_dimReduc>>
```

### Colored by batch {.tabset .tabset-fade}

#### UMAP
```{r rmd_dimreduc_ggplot_covariables}
useReduction='umap'
<<dimreduc_ggplot_covariables>>
rm("useReduction");
```

<!-- #### tSNE -->
<!-- ```{r rmd_heterogeneity_dimReduc_batch_plot_tSNE} -->
<!-- useReduction='tsne'; -->
<!-- <<dimreduc_ggplotly_overlayBatches>> -->
<!-- rm("useReduction"); -->
<!-- ``` -->

<!-- Ending the tabset section -->
<!-- ### {.toc-ignore} -->
<br /><br />




## Clusters identification

### Cells distribution
```{r rmd_heterogeneity_identifyClusters}
<<heterogeneity_identifyClusters>>
```

### Statistics by cluster
```{r rmd_heterogeneity_identifyClusters_splitStats}
<<heterogeneity_identifyClusters_splitStats>>
```
<br /><br />


### Clusters projection {.tabset .tabset-fade}

#### UMAP
```{r rmd_heterogeneity_dimReduc_plot_UMAP}
useReduction='umap'
<<heterogeneity_dimReduc_with_clusters>>
rm("useReduction");
```

<br /><br />


## Marker genes
```{r rmd_heterogeneity_markerGenes}
<<heterogeneity_markerGenes>>
```

### Marker genes list
```{r rmd_heterogeneity_markerGenes_table}
<<heterogeneity_markerGenes_table>>
```
<br />

### Marker genes expression {.tabset .tabset-fade}

#### UMAP {.tabset .tabset-fade}
```{r rmd_heterogeneity_markerGenes_expression_projection_UMAP, fig.dim = c(4, 4), out.width='25%'}
useReduction='umap';

cat("\n<a href='#clusters-projection'>");
<<heterogeneity_dimReduc_thumbnail>>
cat("</a> \n \n");

<<heterogeneity_markerGenes_expression_projection>>

rm("useReduction");
```

<!-- Ending the tabset section -->
<!-- #### {.toc-ignore} -->

<!-- #### tSNE {.tabset .tabset-fade} -->
<!-- ```{r rmd_heterogeneity_markerGenes_expression_projection_tSNE, fig.dim = c(4, 4), out.width='25%'} -->
<!-- useReduction='tsne'; -->

<!-- cat("\n<a href='#clusters-projection'>"); -->
<!-- <<heterogeneity_dimReduc_thumbnail>> -->
<!-- cat("</a> \n \n"); -->

<!-- <<heterogeneity_markerGenes_expression_projection>> -->

<!-- rm("useReduction"); -->
<!-- ``` -->

<!-- <!-- Ending the tabset section --> -->
<!-- #### {.toc-ignore} -->

#### Clusters violin {.tabset .tabset-fade}
```{r rmd_heterogeneity_markerGenes_expression_violin, fig.dim = c(4, 4), out.width='25%'}
cat("\n<a href='#clusters-projection'>");
useReduction='umap'; #c( 'umap', 'tsne');
<<heterogeneity_dimReduc_thumbnail>>
rm("useReduction");
cat("</a> \n \n");

<<heterogeneity_markerGenes_expression_violin>>
```


#### Clusters functional enrichment {.tabset .tabset-fade}
```{r rmd_heterogeneity_markerGenes_functional_enrichment, fig.dim = c(9, 9) }
cat("\n<a href='#clusters-projection'>");
useReduction='umap'; #c( 'umap', 'tsne');
<<heterogeneity_dimReduc_thumbnail>>
rm("useReduction");
cat("</a> \n \n");

<<heterogeneity_markerGenes_functional_enrichment>>
```

#### Cluster heatmap

```{r rmd_heterogeneity_markerGenes_heatmap_mean, fig.dim = c(10, 10)}

<<heterogeneity_markerGenes_heatmap_mean>>

```

<br /><br />




```{r rmd_final_saveSessionImage}
<<final_saveSessionImage>>
```





